---
title: "Local Lion"
author: "Bowen Jones"
date: '`r format(Sys.time(), "%b %d, %Y")`'
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(nycflights13)
```

# Introduction

Here it is! My first real-world experience with data analysis. Up to this point, everything I've done regarding data analysis has been in an educational, guided context. My previous experience has been helpful in providing me with an outline of how to proceed, but in order to fully understand the process I knew I needed true practice. So I spoke with my employer -- the owner of Local Lion, a coffee shop where I've worked since 2013 -- about gaining access to the shop's sales records with the goal of building a model to predict the number of donuts the shop needs to make on a daily basis (to minimize loss). He granted me access to all the sales records since the shop opened in 2012! 
  
I'm super excited to dive in, but I need to solidfy my understanding of data science first. In order to adequately clean and wrangle the data, I plan to begin by covering Chapters 4 and 5 of *Modern Data Science with R*, and completing the relevant exercies below. I also plan on completing the relevant *Data Science* course on DataCamp.

### Chapter 4: Data Wrangling
Exercise 4.1  
Each of these tasks can be performed using a single data verb. For each task, say which verb it is:  
1. Find the average of one of the variables.  
`summarize()`  
2. Add a new column that is the ratio between two variables.  
`mutate()`  
3. Sort the cases in descending order of a variable.  
`arrange()`  
4. Create a new data table that includes only those cases that meet a critereon.  
`filter()`  
5. From a data table with three categorical variables, A, B, and C, and a quantitative variable X, produce a data frame that has the same cases but only the variables A and x.  
`select()`  

Exercises 4.2-3  
Use the `nycflights13` package and the `flights` data frame to answer the following questions: What month had the highest proportion of cancelled flights? What month had the lowest?
```{r}
flights %>%
  group_by(month) %>% 
  summarize(flights = n(), cancelled = sum(is.na(dep_time)), prop = sum(is.na(dep_time))/n()) %>% 
  arrange(prop)
```
October had the lowest proportion of cancelled flights and February had the highest.

What plane (specified by the `tailnum` variable) traveled the most times from New York City airports in 2013? Plot the number of trips per week over the year.
```{r}
flights %>%
  select(flight, tailnum) %>% 
  group_by(tailnum) %>% 
  summarize(flights = n()) %>% 
  arrange(desc(flights))
```

