---
title: "Notes"
author: "Bowen Jones"
date: "9/7/2017"
output: 
  html_document:
    toc: true
    theme: flatly
---
Note: Crtl + alt + I inserts code chunk

# **Binomial Distribution**

Notated as:
$$X~Binom(n, p) $$
Where random variable *X* follows the Binomial Distribution with *n* trials of probability *p*.
To determine *expected value E(X):*

$$E(X) = n * p$$
So, to find the expected value *E(X)* of *X~Binom(10, .5):* 
```{r}
(10*.5)
```
Variance is computed using 
$$Var(X)= E(X) * (1-p)$$

## *Determining the probability of a specific outcome x*
Notated as $Pr(X=x)$, where *x* is the desired number of successes.

### Simulation Method
A common method of determing $Pr(X=x)$ is by running simulations and averaging the results. We use the *rbinom* function as follows: $$rbinom(n, size, p)$$
Note that here, *n* is desired number of simulations, *size* is the number of trials and *p* is the probability of success.
To determine the probability of *x=5* successes given *X~Binom(10, .5)*, 
``` {r}
mean(rbinom(10000, 10, .5)==5)
```
This returns the fraction of outcomes with 5 successes.

### Exact Method
Alternatively, one could use the *dbinom* function to find the exact value as follows, `dbinom(n, size, p)`, where the first argument specifies the desired number of successes. So, 

``` {r}
dbinom(5, 10, .5)
```

mean(rbinom(10000, 10, .5)<=4)
or, pbinom(4, 10, .5)


fair <- dbinom(13, 20, 0.50)
bias <- dbinom(13, 20, 0.80)
fair / (fair + bias)


P(A and B) = P(A)P(B) -- branches
or, 
A <- rbinom(10000, 1, .5)
B <- rbinom(10000, 1, .5)
mean(A & B)

P(A or B) = (P(A) + P(B)) - P(A&B)
mean(A|B)
mean(A <=3 | B <= 3)

multiplying a random variable k results 
in a similar transformation in E(X) -- E|k*X| = k * E(X)
Var(k*Y) = k^2 Var(Y)

fair <- rbinom(10000, 20, .5)
sum(fair==14) = 1888
biased <- rbinom(10000, 20, .75)
sum(biased==14) = 8372
number biased/number total = 8372/(1888+8372)

fair <- dbinom(16, 20, 0.5)
bias <- dbinom(16, 20, 0.8)
p_fair <- 0.99
p_bias <- 0.01
(fair * p_fair) / (fair * p_fair + bias * p_bias)

Var(aX +- bY) = a^2Var(x) + b^2Var(x)
prior prob - alter n

# **Geometric Distribution**
Waiting for first success (heads of flips, in this case; "machine malfunctions" for application)
 
``` {r}
flips <- rbinom(100, 1, .1)
```
Using the *which* function, 
``` {r} 
which(flips == 1)[1] 
```

For simulation method use the *replicate* function, yielding a list of the first appearance of a success for each of 10 trials 

``` {r} 
replicate(10, which(rbinom(100, 1, .1)==1)[1])
```
To determine expected value, run mean
``` {r} 
mean(sum(replicate(10, which(rbinom(100, 1, .1)==1)[1])))
```

##Introducing rgeom
*style:* rgeom(draws, probability)

X~Geom(*p*)
``` {r}
geom <- rgeom(100000, .1)
mean(geom)
```
$E(X) = (1/p) - 1$