---
title: "R Graphics Cookbook Chapter 9"
author: "Bowen W. Jones"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
```

## Controlling the Overall Appearance of Graphs

Note: Some problems cannot be completed as written because the code has changed. You will need to find the updated code.

### Section 9.1. Setting the Title of a Graph

```{r}
# Figure 9-1
p <- ggplot(heightweight, aes(ageYear, heightIn)) + geom_point()
p + ggtitle("Age and Height of Schoolchildren")
p + ggtitle("Age and Height\nof Schoolchildren")

# Figure 9-2
p + ggtitle("Age and Height of Schoolchildren") +
  theme(plot.title=element_text(vjust = -6.5))
p + annotate("text", x = mean(range(heightweight$ageYear)), y = Inf,
             label = "Age and Height of Schoolchildren", vjust = 1.5, size = 6)
```

### Section 9.2. Changing the Appearance of Text

```{r}
# Figure 9-3
p + theme(axis.title.x = element_text(size = 16, lineheight = 0.9, family = "Times", face = "bold.italic", color = "red"))
p + ggtitle("Age and Height\nof Schoolchildren") +
    theme(axis.title.x = element_text(size = rel(1.5), lineheight = 0.9, family = "Times", face = "bold.italic", color = "red"))
p + annotate("text", x = 15, y = 53, label = "Some text", size = 7, family = "Times", fontface = "bold.italic", color = "red")
p + geom_text(aes(label = weightLb), size = 4, family = "Times", color = "red")
```

### Section 9.3. Using Themes

```{r}
# Figure 9-5
p <- ggplot(heightweight, aes(ageYear, heightIn, color = sex)) + geom_point()
p + theme_grey()
p + theme_bw()
p + theme_grey(base_size = 16, base_family = "Times")
theme_set(theme_bw())
p
theme_set(theme_grey())
```

### Section 9.4. Changing the Appearance of Theme Elements

```{r}
# Figure 9-6 
p + theme(
  panel.grid.major = element_line(color = "red"),
  panel.grid.minor = element_line(color = "red", linetype = "dashed", size = 0.2),
  panel.background = element_rect(fill = "lightblue"),
  panel.border = element_rect(color = "blue", fill = NA, size = 2))
p + ggtitle("Plot title here") + 
  theme(
  axis.title.x = element_text(color = "red", size = 14),
  axis.text.x = element_text(color = "blue"),
  axis.title.y = element_text(color = "red", size = 14, angle = 90),
  axis.text.y = element_text(color = "blue"),
  plot.title = element_text(color = "red", size = 20, face = "bold"))
p + theme(
  legend.background = element_rect(fill = "grey85", color = "red", size = 1),
  legend.title = element_text(color = "blue", face = "bold", size = 14),
  legend.text = element_text(color = "red"),
  legend.key = element_rect(color = "blue", size = 0.25))
p + facet_grid(sex ~ .) + theme(
  strip.background = element_rect(fill = "pink"),
  strip.text.y = element_text(size = 14, angle = -90, face = "bold")
)

```

### Section 9.5. Creating Your Own Themes

```{r}
# Figure 9-7
mytheme <- theme_bw() +
  theme(text = element_text(color = "red"),
        axis.title = element_text(size = rel(1.25)))
p <- ggplot(heightweight, aes(ageYear, heightIn)) + geom_point()
p + mytheme
```

### Section 9.6. Hiding Grid Lines

```{r}
# Figure 9-8
p + theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
p + theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())
p + theme(panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())
```

<hr>
END!