---
title: "R Graphics Cookbook Chapter 5"
author: "Bowen W. Jones"
date: "`r Sys.Date()`"
output: html_document
---

### Chapter 5. Scatter Plots

```{r}
library(ggplot2)
library(gcookbook)
```

### Section 5.1. Making a Basic Scatterplot

```{r}
# Figure 5-1
ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point()

# Figure 5-2
ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point(shape = 21)
ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point(size = 1.5)

# Figure 5-3
ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point(shape = 16)
ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point(shape = 19)

```

### Section 5.2. Grouping Data Points by a Variable Using Shape or Color

```{r}
# Figure 5-4
ggplot(heightweight, aes(ageYear, heightIn, color = sex)) +
  geom_point()
ggplot(heightweight, aes(ageYear, heightIn, shape = sex)) +
  geom_point()

# Figure 5-5
ggplot(heightweight, aes(ageYear, heightIn, shape = sex, color = sex)) +
  geom_point()
ggplot(heightweight, aes(ageYear, heightIn, shape = sex)) +
  geom_point() +
  scale_shape_manual(values=c(1,2)) +
  scale_color_brewer(palette="Set1")

```

### Section 5.3. Using Different Point Shapes

```{r}
# Figure 5-6
ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point(shape = 3)
ggplot(heightweight, aes(ageYear, heightIn, shape = sex)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c(1, 4))

# Figure 5-8
hw <- heightweight
hw$weightGroup <- cut(hw$weightLb, breaks = c(-Inf, 100, Inf),
                      labels = c("< 100", ">= 100"))
ggplot(hw, aes(ageYear, heightIn, shape = sex, fill = weightGroup)) +
  geom_point(size = 2.5) +
  scale_shape_manual(values = c(21, 24)) +
  scale_fill_manual(values = c(NA, "black"),
                    guide = guide_legend(override.aes = list(shape = 21)))
```

### Section 5.4. Mapping a Continuous Variable to Color or Size

```{r}
# Figure 5-9
ggplot(heightweight, aes(ageYear, heightIn, color = weightLb)) +
  geom_point()
ggplot(heightweight, aes(ageYear, heightIn, size = weightLb)) +
  geom_point()

# Figure 5-10

ggplot(heightweight, aes(weightLb, heightIn, fill = ageYear)) +
  geom_point(shape = 21, size = 2.5) +
  scale_fill_gradient(low = "black", high = "white")
ggplot(heightweight, aes(weightLb, heightIn, fill = ageYear)) +
  geom_point(shape = 21, size = 2.5) +
  scale_fill_gradient(low = "black", high = "white", breaks = 12:17,
                      guide = guide_legend())
# Figure 5-11
ggplot(heightweight, aes(ageYear, heightIn, size = weightLb, color = sex)) +
  geom_point(alpha = 0.5) +
  scale_size_area() +
  scale_color_brewer(palette = "Set1")
```

### Section 5.5 Dealing with Overplotting

```{r}
# Figure 5-12
sp <- ggplot(diamonds, aes(carat, price))
sp + geom_point()

# Figure 5-13
sp + geom_point(alpha = 0.1)
sp + geom_point(alpha = 0.01)

# Figure 5-14
sp + stat_bin_2d()
sp + stat_bin_2d(bins = 50) +
  scale_fill_gradient(low = "lightblue", high = "red", limits = c(0, 6000))

# Figure 5-15
library(hexbin)
sp + stat_binhex() +
  scale_fill_gradient(low = "lightblue", high = "red", limits = c(0, 8000))
sp + stat_binhex() +
  scale_fill_gradient(low = "lightblue", high = "red", breaks  = c(0,250,500,1000,2000,4000,6000),   limits = c(0, 6000))

# Figure 5-16
sp1 <- ggplot(ChickWeight, aes(Time, weight))
sp1 + geom_point()
sp1 + geom_point(position = "jitter")
sp1 + geom_point(position = position_jitter(width = 0.5, height = 0))

# Figure 5-17
sp1 + geom_boxplot(aes(group = Time))

```

### Section 5.6. Adding Fitted Regression Model Lines

```{r}
# Figure 5-18
sp <- ggplot(heightweight, aes(ageYear, heightIn))
sp + geom_point() +
  stat_smooth(method = lm)
sp + geom_point() +
  stat_smooth(method = lm, level = 0.99)
sp + geom_point() +
  stat_smooth(method = lm, se = FALSE)

# Figure 5-19
sp + geom_point(color = "grey60") +
  stat_smooth()
sp + geom_point(color = "grey60") +
  stat_smooth(method = loess)

# Figure 5-20
library(MASS)
b <- biopsy
b$classn[b$class=="benign"] <- 0
b$classn[b$class=="malignant"] <- 1
ggplot(b, aes(V1, classn)) +
  geom_point(position = position_jitter(width = 0.3, height = 0.06), alpha = 0.4,
             shape = 21, size = 1.5) +
  stat_smooth(method = glm, family = binomial)

# Figure 5-21
sps <- ggplot(heightweight, aes(ageYear, heightIn, color = sex)) +
  geom_point() +
  scale_color_brewer(palette = "Set1")
sps + geom_smooth()
sps + geom_smooth(method = lm, se = FALSE, fullrange = TRUE)
```

### Section 5.7. Adding Fitted Lines from an Existing Model

```{r}
# Figure 5-22
model <- lm(heightIn ~ ageYear + I(ageYear^2), data = heightweight)
xmin <- min(heightweight$ageYear)
xmax <- max(heightweight$ageYear)
predicted <- data.frame(ageYear = seq(xmin, xmax, length.out = 100))
predicted$heightIn <- predict(model, predicted)
sp <- ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point(color = "grey40")
sp + geom_line(data = predicted, size = 1)

# Predict values of yvar from xvar given a model

predictvals <- function(model, xvar, yvar, xrange = NULL, samples = 100, ...) {
  if(is.null(xrange)) {
    if(any(class(model) %in% c("lm", "glm")))
      xrange <- range(model$model[[xvar]])
    else if (any(class(model) %in% "loess"))
      xrange <- range(model$x)
  }
  
  newdata <- data.frame(x = seq(xrange[1], xrange[2], length.out = samples))
  names(newdata) <- xvar
  newdata[[yvar]] <- predict(model, newdata = newdata, ...)
  newdata
}

modlinear <- lm(heightIn ~ ageYear, heightweight)
modloess <- loess(heightIn ~ ageYear, heightweight)
lm_predicted <- predictvals(modlinear, "ageYear", "heightIn")
loess_predicted <- predictvals(modloess, "ageYear", "heightIn")
sp + geom_line(data = lm_predicted, color = "red", size = 0.8) +
  geom_line(data = loess_predicted, color = "blue", size = 0.8)

# Figure 5-23
library(MASS)
b <- biopsy

b$classn[b$class=="benign"] <- 0
b$classn[b$class=="malignant"] <- 1

# Perform logistic regression
fitlogistic <- glm(classn ~ V1, b, family = binomial)

# Get predicted values
glm_predicted <- predictvals(fitlogistic, "V1", "classn", type = "response")

ggplot(b, aes(V1, classn)) +
  geom_point(position=position_jitter(width=.3, height=0.8), alpha = 0.4,
             shape = 21, size = 1.5) +
  geom_line(data=glm_predicted, color = "#1177FF", size = 1)
```

### Section 5.8. Adding Fitted Lines from Multiple Existing Models

```{r}
# Figure 5-24
library(plyr)
make_model <- function(data) {
  lm(heightIn ~ ageYear, data)
}
models <- dlply(heightweight, "sex", .fun=make_model)
predvals <- ldply(models, .fun = predictvals, xvar = "ageYear", yvar= "heightIn")

ggplot(heightweight, aes(ageYear, heightIn, color = sex)) +
  geom_point() + geom_line(data = predvals)
ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point() + geom_line(data = predvals) + facet_grid(~ sex)

# Figure 5-25
predvals <- ldply(models, .fun = predictvals, xvar = "ageYear", yvar= "heightIn", 
                  xrange = range(heightweight$ageYear))
ggplot(heightweight, aes(ageYear, heightIn, color = sex)) +
  geom_point() + geom_line(data = predvals)
```

### Section 5.9. Adding Annotations with Model Coefficients

```{r}
# Figure 5-26
model <- lm(heightIn ~ ageYear, heightweight)
# Generate prediction data
pred <- predictvals(model, "ageYear", "heightIn")
sp <- ggplot(heightweight, aes(ageYear, heightIn)) +
  geom_point() + geom_line(data = pred)
sp + annotate("text", label = "r^2 = 0.42", x = 16.5, y = 52)
sp + annotate("text", label = "r^2 == 0.42", parse = TRUE, x = 16.5, y = 52)

# Figure 5-27
eqn <- as.character(as.expression(
  substitute(italic(y) == a + b * italic(x) * "," ~~  italic(r)^2 ~ "=" ~ r2,
             list(a = format(coef(model)[1], digits = 3),
                  b = format(coef(model)[2], digits = 3),
                  r2 = format(summary(model)$r.squared, digits = 2)
                  ))))
sp + annotate("text", label = eqn, parse = TRUE, x = Inf, y = -Inf, hjust = 1.1, vjust = -0.5)
```

### Section 5.10. Adding Marginal Rugs to a Scatterplot

```{r}
# Figure 5-28
ggplot(faithful, aes(eruptions, waiting)) +
  geom_point() +
  geom_rug()

# Figure 5-29
ggplot(faithful, aes(eruptions, waiting)) +
  geom_point() +
  geom_rug(position = "jitter", size = 0.2)

```

### Section 5.11. Labeling Point in a Scatterplot

```{r}
# Figure 5-30
sp <- ggplot(subset(countries, Year==2009 & healthexp>2000), aes(healthexp, infmortality)) + geom_point()
sp + annotate("text", x = 4350, y = 5.4, label = "Canada") +
  annotate("text", x = 7400, y = 6.8, label = "USA")
sp + geom_text(aes(label = Name), size = 4)

# Figure 5-31
sp + geom_text(aes(label = Name), size = 4, vjust = 0)
sp + geom_text(aes(y = infmortality + 0.1, label = Name), size = 4, vjust = 0)

# Figure 5-32
sp + geom_text(aes(label = Name), size = 4, hjust = 0)
sp + geom_text(aes(x = healthexp+100, label = Name), size = 4, vjust = 0)

# Figure 5-33
cdat <- subset(countries, Year==2009 & healthexp>2000)
cdat$Name1 <- cdat$Name
idx <- cdat$Name1 %in% c("Canada", "Ireland", "United Kingdom", "United States", "New Zealand", "Iceland", "Japan", "Luxembourg", "Netherlands", "Switzerland")
cdat$Name1[!idx] <- NA
ggplot(cdat, aes(healthexp, infmortality)) +
  geom_point() +
  geom_text(aes(x = healthexp+100, label = Name1), size = 4, hjust = 0) +
  xlim(2000, 10000)
```

### Section 5.12. Creating a Balloon Plot

```{r}
# Figure 5-34
cdat <- subset(countries, Year == 2009 & 
                 Name %in% c("Canada", "Ireland", "United Kingdom", "United States", "New Zealand", "Iceland", "Japan", "Luxembourg", "Netherlands", "Switzerland"))
p <- ggplot(cdat, aes(healthexp, infmortality, size = GDP)) +
  geom_point(shape = 21, color = "black", fill = "cornsilk")
p
p + scale_size_area(max_size = 15)

# Figure 5-35
hec <- HairEyeColor[,,"Male"] + HairEyeColor[,,"Female"]
library(reshape2)
hec <- melt(hec, value.name = "count")

ggplot(hec, aes(Eye, Hair)) +
  geom_point(aes(size = count), shape = 21, color = "black", fill = "cornsilk") +
  scale_size_area(max_size = 20, guide = FALSE) +
  geom_text(aes(y = as.numeric(Hair)-sqrt(count)/22, label = count), vjust = 1, color = "grey60", size = 4)
```

### Section 5.13. Making a Scatterplot Matrix

```{r}
# Figure 5-36
c2009 <- subset(countries, Year == 2009,
                select = c(Name, GDP, laborrate, healthexp, infmortality))
pairs(c2009[,2:5])

# Figure 5-37
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r <- abs(cor(x, y, use = "complete.obs"))
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste(prefix, txt, sep = "")
  if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
  text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

panel.hist <- function(x, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5))
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks
  nB <- length(breaks)
  y <- h$counts
  y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, col = "white", ...)
}
  
pairs(c2009[,2:5], upper.panel = panel.cor,
              diag.panel = panel.hist,
              lower.panel = panel.smooth)
# Figure 5-38
panel.lm <- function (x, y, col = par("col"), bg = NA, pch = par("pch"),
                      cex = 1, col.smooth = "black",...) {
  points(x, y, pch = pch, col = col, bg = bg, cex = cex)
  abline(stats::lm(y~x), col = col.smooth, ...)
}
pairs(c2009[,2:5], pch = ".",
              upper.panel = panel.cor,
              diag.panel = panel.hist,
              lower.panel = panel.lm)
```


<hr>
END!